---
title: "R Notebook"
output: html_notebook
---

# Data cleaning
_Attempts to standardize columns into the same format and seperate out depth and country information_

## dependancies
```{r dependancies}
library(tidyverse)
library(here)
library(leaflet)
library(readxl)
library(janitor)
library(googlesheets)
#source("./helpers_masterlist.R")
library(revgeo)



```

download google sheet
```{r}
## Global

mysheets <- gs_ls()

gs_master <- gs_title("PTM - Master List")

#gets the data
gs_ptm<- gs_read(gs_master)

#cleans up columns
colnames(gs_ptm) <- gs_ptm[1,]
new <- gs_ptm[-c(1,2),]
```

# GPS Coordinates
- Cleaning up all that are not decimal degrees
- non decimal degree coordinates are less machine readable (ie DD"MM'SS format may mutate if the file is saved differently)
```{r echo=TRUE}
#all records that have decimal degrees
ptm_ll <- filter(new,
                 !is.na(Latitude),
                 !str_detect(new$Latitude,"\""),
                 !str_detect(new$Latitude,"\'"),
                 !str_detect(new$Latitude,"N"))
```

## Mapping the data in leaflet reveals the issues
- coordinates are poor for Chile (- are not in latitude or longitude) -> plotted in Kyrzkstan
### plotting data to check work
```{r echo=FALSE}
#function for mapping
mapping <- function(ds) {
  mn <- ds %>% 
  mutate(Latitude = as.numeric(Latitude),
         Longitude = as.numeric(Longitude))
  
  map <- leaflet() %>% 
  addProviderTiles("CartoDB") %>% 
  addCircleMarkers(lng = mn$Longitude, lat = mn$Latitude,
             popup = paste(mn$Locality, mn$`PTM#`, "Lat",mn$Latitude, "Long", mn$Longitude),
             clusterOptions = markerClusterOptions())
  return(map)
}

mapping(ptm_ll)

# gets the Country and State information from the Longitude and Latitude information
test <- tail(ptm_ll)
rev <- revgeo(ptm_ll$Longitude,ptm_ll$Latitude,output = "frame")
alarm()
```


### transforming data to match
* maybe you can use a for loop here
```{r}
change <- c("island","Island")

for (x in change) {
  ptm_ll$Longitude = ifelse(grepl(x,ptm_ll$Locality) & !grepl("-",ptm_ll$Longitude),
  paste0("-",ptm_ll$Longitude),
  ptm_ll$Longitude)
}

ptm_ll$Longitude = ifelse(grepl("Chile",ptm_ll$Locality) & !grepl("-",ptm_ll$Latitude),
  paste0("-",ptm_ll$Longitude),
  ptm_ll$Longitude)

ptm_ll$Latitude = ifelse(grepl("Chile",ptm_ll$Locality) & !grepl("-",ptm_ll$Latitude),
  paste0("-",ptm_ll$Latitude),
  ptm_ll$Latitude)
```

### plotting data again to check work
```{r echo=FALSE}
cleaned <- mapping(ptm_ll)
```

convert to decimal degrees
- need to resolve: 1718 (in NT?)
- converts notation (ie. " ') to spaces to be passed to conv_unit (format accepted)
- careful! needed to add negative sign to make sure W values are properly converted
```{r}
#all records that need cleaning
ptm_cord <- filter(new,
                 !is.na(Latitude),
                 !new$`PTM#` %in% ptm_ll$`PTM#`)

# change the symbols to a space
symbol <- c('°','\'')
for (y in symbol) {
  ptm_cord$Latitude = gsub(y, ' ', ptm_cord$Latitude)
}

char <- c('\"',"\\")
for (c in char) {
  ptm_cord$Latitude = gsub(char, "", ptm_cord$Latitude)
}
ptm_cord$Latitude = gsub('N', '', ptm_cord$Latitude)
ptm_cord$Latitude = trimws(ptm_cord$Latitude, which = "both")

ptm_cord$Longitude = ifelse(grepl("W",ptm_cord$Longitude),
  paste0("-",ptm_cord$Longitude),
  ptm_cord$Longitude)

lc <- c('W','E', '\"')
for(c in lc) {
  ptm_cord$Longitude = gsub(c, '', ptm_cord$Longitude)
}

for (x in symbol) {
  ptm_cord$Longitude = gsub(x, ' ', ptm_cord$Longitude)
}

ptm_cord$Longitude = trimws(ptm_cord$Longitude, which = "both")
# convert from decimal minutes to decimal degrees
ptm_cord$Latitude = measurements::conv_unit(ptm_cord$Latitude, from = 'deg_min_sec', to = 'dec_deg')
ptm_cord$Longitude = measurements::conv_unit(ptm_cord$Longitude, from = 'deg_min_sec', to = 'dec_deg')
```

# checking converted data
```{r}
 mapping(ptm_cord)

## up to here
```

putting back the data together and plotting
- come back and look why ptm_gps does not equal to 1665
```{r}
new_na <- filter(new,
                 is.na(Latitude))
ptm_gps <- rbind(ptm_ll, ptm_cord, new_na)

mapping(ptm_gps)
```

## Date collected
```{r}
group_by(ptm_gps, `Date Collected`) %>% 
  summarise()
ptm_gps <- ptm_gps %>%
  mutate(`Date Collected` = as.numeric(`Date Collected`))
xls_date <-  ptm_gps %>% 
  mutate (converted = excel_numeric_to_date(ptm_gps$`Date Collected`,
                                                 date_system = "modern"))
```

## Final Determination
- cleaning up names
```{r}
species <- ptm_gps %>% count(`Final determination`, sort = TRUE)
ptm_gps$`Final determination` <- gsub("Chondrachanthus", "Chondracanthus", ptm_gps$`Final determination`)

ptm_gps$`Final determination` <- gsub("\r\n", "", ptm_gps$`Final determination`)
ptm_gps$`Final determination` <- gsub("Sp\\.", "sp.", ptm_gps$`Final determination`)
ptm_gps$`Final determination` <- gsub("sp. 2", "sp.2", ptm_gps$`Final determination`)
ptm_gps$`Final determination` <- gsub("unknown species", "unknown", ptm_gps$`Final determination`)
```

## Coralline/ Red/ Green / Brown
ensuring uniform data
```{r eval=FALSE, include=FALSE}
group_by(ptm_gps, `Red/Coralline/Green/Brown`) %>% 
  summarise()
ptm_gps$`Red/Coralline/Green/Brown` <- gsub("red", "Red", ptm_gps$`Red/Coralline/Green/Brown`)
ptm_gps$`Red/Coralline/Green/Brown` <- gsub("\r\n", "", ptm_gps$`Red/Coralline/Green/Brown`)
  
```

## correcting for inconsistencies
### Main Collector and Determiners
using gsub to fix
```{r}
group_by(ptm_gps, `Main Collector`) %>% 
  summarise()
ptm_gps$`Main Collector` <- gsub("P.T.Martone", "Patrick T. Martone", ptm_gps$`Main Collector`) 
ptm_gps$`Main Collector` <- gsub("PTMartone", "Patrick T. Martone", ptm_gps$`Main Collector`)
ptm_gps$`Main Collector` <- gsub("Samuel", "Sam", ptm_gps$`Main Collector`)
ptm_gps$`Main Collector` <- gsub("Vanmaanen", "VanMaanen", ptm_gps$`Main Collector`) 
group_by(ptm_gps, `Other collectors`) %>% 
  summarise()
group_by(ptm_gps, `Determiners`) %>% 
  summarise()
ptm_gps$`Determiners` <- gsub("Samuel", "Sam", ptm_gps$`Determiners`)
group_by(ptm_gps, `Fragments available?`) %>% 
  summarise()
ptm_gps$`Fragments available?` <- gsub("yes", "Yes", ptm_gps$`Fragments available?`)
ptm_gps$`Fragments available?` <- gsub("y", "Yes", ptm_gps$`Fragments available?`)
```

## Habitat and Locality
extract depth data, accounts for # m, #m, # ft, #ft
need to account for -0.5 - 0.0m
easy way to include 3 digit numbers? or with decimals ##.#
probably use a(?=c) a followed by c some how
```{r echo=TRUE}
group_by(ptm_gps, `Habitat`) %>% 
  summarise()
ptm_gps$Habitat <- gsub("Subtidal", "subtidal", ptm_gps$Habitat)
group_by(ptm_gps, `Missing accession`) %>% 
  summarise()
depth <- ptm_gps %>%
  filter(grepl("subtidal",Habitat),
         grepl(",", Habitat)) %>% 
  separate(col = Habitat, into = c("Habitat", "Depth", "Other"), sep = ",")
#Locality
point <- str_replace(ptm_gps$Locality,"Pt|Pt.","Point")
#countries
group_by(ptm_gps, `PCR date (CO1)`) %>% 
  summarise()
```

```{r}
ptm_gps <- ptm_gps %>% 
  mutate(Latitude = as.numeric(Latitude),
         Longitude = as.numeric(Longitude),
         `PTM#` = as.numeric(ptm_gps$`PTM#`))

write_csv(ptm_gps, "ptm_clean.csv")
```

