---
title: "01 Filling Bold Spreadsheet"
author: "Jasmine"
date: '2019-03-01'
output: html_document
---

This notebook will walk you through how to use R to quickly take the PTM master list
# how to use
- this code is previously configured to work with an example (run the code chunks to see the output)
- make a copy of this folder and edit the code to work for your dataset

make sure if you have the dependancies:
- if not type in `install.packages()` to your console with the package name

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(openxlsx)
library(taxize)
library(here)
```

# Loading the data

## PTM master list

1. make sure you have the most recent version of the file saved from the google drive
2. place the file in the 

Google sheet documentation:
https://cran.r-project.org/web/packages/googlesheets/vignettes/basic-usage.html
```{r}
library(googlesheets)
#library(googlesheets4) # future will be googlesheets4
#(ptm <- drive_get("PTM - Master List"))

#need to authenticate (see popup window)
mysheets <- gs_ls()

gs_master <- gs_title("PTM - Master List")

#gets the data
gs_ptm<- gs_read(gs_master, ws = 1) 

#cleans up columns
colnames(gs_ptm) <- gs_ptm[1,]
new <- gs_ptm[-c(1,2),]
```

### other collectors
```{r}
#gets the data
gs_other<- gs_read(gs_master, ws = 3)

#cleans up columns
colnames(gs_other) <- gs_other[1,]
new_other <- gs_other[-c(1,2),]
```


If you don't want to use the google sheet api
```{r}
#master <- read_csv(here("01_spreadsheet/examples/PTM_Master.csv"))
```

## Data of interest
1. put the file with all the PTM numbers you want in this folder
2. replace the example "Quad_Data.csv"
```{r}
want <- read_csv(here("01_spreadsheet/examples/BOLD_2019Jul25.csv"))

want_2 <- want %>% 
  filter(!is.na(`PTM#`)) %>%
  mutate(`PTM#` = str_replace(`PTM#`,"PTM",""))
```

# filtering for the PTM# you want to submit and removing empty cells
```{r}
submit_ptm <- new %>% 
  filter(new$`PTM#` %in% want_2$`PTM#`,
         !is.na(`PTM#`))

submit_other <- new_other %>% 
  filter(new_other$`PTM#` %in% want_2$`PTM#`)

submit <- rbind(submit_ptm,submit_other)
```

## Get the Bold Template
```{r}
info <- read_xls("SpecimenDataV3Transitional_BOLDTemplate.xls", skip = 1, sheet = 1)

e_info <- info[FALSE,]

tax <- read_xls("SpecimenDataV3Transitional_BOLDTemplate.xls", skip = 1, sheet = 2)

e_taxon <- tax[FALSE,]

spe <- read_xls("SpecimenDataV3Transitional_BOLDTemplate.xls", skip = 1, sheet = 3)

e_specimen <- spe[FALSE,]

coll <- read_xls("SpecimenDataV3Transitional_BOLDTemplate.xls", skip = 1, sheet = 4)

e_collect <- coll[FALSE,]
```

# Higher Taxonomy

get the species you want
*MOD - May need to combine between the Final determination and the field identification columns

* note there is a limit to the number of entries you can retrieve at one time so subset the dataframe if needed

- change depending on field `Final Determination` or `Determination in the field`
```{r}
df_nm <- submit %>% 
  separate(col = `Determination in the field`,
         into = c("genus","species","other"),
         sep = "[:blank:]",
         extra = "drop") %>% 
  mutate(species = str_replace(species,"\\?",""),
         name = paste(genus,species))

higher <- classification(df_nm$name, db="bold")
```

Get data for each species and write into a table
- avoids NA
- writes to a new dataframe
- fills in Sample ID
```{r}
for(i in 1:length(higher)) {
  if(!is.na(higher[i])) {
    species <- higher[[i]]
     rect <- c(species$name)
     names(rect) <- c(species$rank)
     
    e_taxon <- e_taxon %>%  add_row(
      Phylum = rect["phylum"],
      Order = rect["order"],
      Family = rect["family"],
      Subfamily = rect["subfamily"],
      Species = rect["species"],
      Class = rect["class"],
      Genus = rect["genus"])
  }
  else{e_taxon <- e_taxon %>%  add_row(Phylum = NA)}
}
higher_tax <- e_taxon %>% 
  mutate(`Sample ID` = submit$`PTM#`)

e_taxon <- tax[FALSE,]
```

## Voucher Sheet
** CAUTION CHECK THROUGH THIS PART**
Edit the data to match the formatting of BOLD data
```{r}
c_info <- submit %>% 
  mutate(# Voucher
    Sample_ID = ifelse(str_detect(`PTM#`,"^[:alpha:]"),
                       `PTM#`,
                       paste0("PTM",`PTM#`)),
    Field_ID = Sample_ID,
    Museum_ID = `Accession #`,
    Collection_Code = "",
    Insitution_Storing = "University of British Columbia, Herbarium",
    #Taxon
    Species = `Determination in the field`,
    Identifier = `Determiners`,
    Identifier_Email = "", # !!! make this correspond
    # pmartone@mail.ubc.ca, samuel.starko@gmail.com
    Identifier_Insitution  = "University of British Columbia",
    `Identification_Method` = "",
    `Taxonomy Notes` = "", # !!! make this an available field
    #collect
    Collectors = "",
    Collection_Date = `Date Collected`,
    Country_Ocean = Country,
    State_Province = StateProvince,
    Region = "",
    #Specimen Details
    Sex = `Reproductive Status`,
    Reproduction = "",
    Life_Stage = "",
    Extra_info = "",
    Note = "",
    Voucher_stat = "",
    Tissue_des = "",
    Assoc_tax = "",
    Assoc_spe = "",
    External = ""
    ) %>% #depends on dataset,
  separate(Notes, c("Site_Code","Notes"), ";", extra = "drop") %>%
  separate(Locality, c("Sector","Exact_site"),",")

c_info_2 <- c_info %>% 
  mutate(Sector = str_trim(Sector),
         Exact_site = str_trim(Exact_site),
         Elevation = "",
         Depth = Depth,
         Elevation_Precision = "",
         Depth_Precision = "",
         GPS_Source	= "",
         Coordinate_Accuracy= "",
         Event_Time = "",
         Collection_Date_Accuracy	= "",
         Habitat = Habitat,
         Sampling_Protocol = "",
         Collection_Notes = Notes,
         Site_Code = Site_Code, # !!! make this an available field
         Collection_Event_ID = "")
```

```{r}
#selecting the relevant columns
voucher <- c_info %>%
  select(`Sample_ID`,`Field_ID`, `Museum_ID`, `Collection_Code`, `Insitution_Storing`)

names(voucher) <- names(e_info)
```

```{r}
s <- c_info_2 %>% 
  select(`Sample_ID`)

t_high <-  higher_tax %>%
  select(`Phylum`, `Class`, `Order`, `Family`, `Subfamily`, `Genus`)
t_sub <- c_info %>% 
  select(`Species`,`Identifier`, `Identifier_Email`, `Identifier_Insitution`, `Identification_Method`, `Taxonomy Notes`)

taxon <- cbind(s,t_high,t_sub)
names(taxon) <- names(e_taxon)

collect <-  c_info_2 %>%
  select(`Sample_ID`,`Collectors`, `Collection_Date`, `Country_Ocean`,`State_Province`, `Region`, `Sector`, `Exact_site`, `Latitude`, `Longitude`,`Elevation`, Depth, Elevation_Precision, Depth_Precision, Depth_Precision, GPS_Source, Coordinate_Accuracy, Event_Time, Collection_Date_Accuracy, Habitat, Sampling_Protocol, Collection_Notes, Site_Code, Collection_Event_ID)

names(collect) <- names(e_collect)
```

```{r}
sd <- c_info %>% 
  select(Sample_ID,Sex, Reproduction, Life_Stage, Extra_info, Note, Voucher_stat, Tissue_des, Assoc_tax, Assoc_spe, External)

names(sd) <- names(e_specimen)
```

# Finally saving your data
- always look through the final records to see if the information has been placed correctly
```{r}
lod <- list("Voucher Info" = voucher, "Taxonomy" = taxon, "Specimen Details" =sd, "Collection Data" = collect )
write.xlsx(lod, file = "results/bold_jul262019.xlsx", append = TRUE)
```
