---
title: "01 Filling Bold Spreadsheet"
author: "Jasmine"
date: '2019-03-01'
output: html_document
---

This notebook will walk you through how to use R to quickly take the PTM master list
# how to use
- this code is previously configured to work with an example (run the code chunks to see the output)
- make a copy of this folder and edit the code to work for your dataset

make sure if you have the dependancies:
- if not type in `install.packages()` to your console with the package name

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(openxlsx)
library(taxize)
library(here)
library(googlesheets)
library(beepr)
```

## Data of interest
1. put the file with all the PTM numbers you want in this folder
2. replace the example "Quad_Data.csv"
```{r}
phyla <- c("Red","Coralline")
want <- "bioblitz2019"
filename <- paste0("bioblitz_2019_",phyla)
```

# Loading the data

## PTM master list

1. make sure you have the most recent version of the file saved from the google drive
2. place the file in the 

Google sheet documentation:
https://cran.r-project.org/web/packages/googlesheets/vignettes/basic-usage.html

!!! eventually change this to googlesheets4 and googledrive
```{r}
#library(googlesheets4) # future will be googlesheets4
#(ptm <- drive_get("PTM - Master List"))

#need to authenticate (see popup window)
mysheets <- gs_ls()

gs_master <- gs_title("PTM - Master List")

#gets the data
gs_ptm<- gs_read(gs_master)

#cleans up columns
colnames(gs_ptm) <- gs_ptm[1,]
new <- gs_ptm[-c(1,2),]
```

# filtering for the PTM# you want to submit and removing empty cells
# note to self - update 1882 to Coralline
```{r}
new %>% 
  filter(Project == want) %>% 
  group_by(`Red/Coralline/Green/Brown`) %>% 
  summarise(n()) 

new %>% 
  filter(Project == want) %>% 
  nrow()

#1894 1936

submit <- phyla %>% 
  map_dfr(~filter(new, Project == want, 
                  `Red/Coralline/Green/Brown` == .))
```

## Get the Bold Template
```{r}
info <- read_xls("SpecimenDataV3Transitional_BOLDTemplate.xls", skip = 1, sheet = 1)

e_info <- info[FALSE,]

tax <- read_xls("SpecimenDataV3Transitional_BOLDTemplate.xls", skip = 1, sheet = 2)

e_taxon <- tax[FALSE,]

spe <- read_xls("SpecimenDataV3Transitional_BOLDTemplate.xls", skip = 1, sheet = 3)

e_specimen <- spe[FALSE,]

coll <- read_xls("SpecimenDataV3Transitional_BOLDTemplate.xls", skip = 1, sheet = 4)

e_collect <- coll[FALSE,]
```

# Higher Taxonomy

get the species you want
*MOD - May need to combine between the Final determination and the field identification columns

* note there is a limit to the number of entries you can retrieve at one time so subset the dataframe if needed
```{r}
higher <- classification(submit$`Determination in the field`, db="bold")
```

Get data for each species and write into a table
- avoids NA
- writes to a new dataframe
- fills in Sample ID
```{r}
for(i in 1:length(higher)) {
  if(!is.na(higher[i])) {
    species <- higher[[i]]
     rect <- c(species$name)
     names(rect) <- c(species$rank)
     
    e_taxon <- e_taxon %>%  add_row(
      Phylum = rect["phylum"],
      Order = rect["order"],
      Family = rect["family"],
      Subfamily = rect["subfamily"],
      Species = rect["species"],
      Class = rect["class"],
      Genus = rect["genus"])
  }
  else{e_taxon <- e_taxon %>%  add_row(Phylum = NA)}
}
higher_tax <- e_taxon %>% 
  mutate(`Sample ID` = submit$`PTM#`)

e_taxon <- tax[FALSE,]
```

```{r}
# 
# df <- submit %>% 
#   mutate(taxonomy =  classification(submit$`Determination in the field`, db="bold"))
# 
# test <- df %>% 
#   mutate(species = tax_name(query = df$`Determination in the field`,
#                             get = "species",
#                             db = "ncbi"))
# 
# for(i in 1:length(higher)) {
#   if(!is.na(higher[i])) {
#     species <- higher[[i]]
#      rect <- c(species$name)
#      names(rect) <- c(species$rank)
#      
#     e_taxon <- e_taxon %>%  add_row(
#       Phylum = rect["phylum"],
#       Order = rect["order"],
#       Family = rect["family"],
#       Subfamily = rect["subfamily"],
#       Species = rect["species"],
#       Class = rect["class"],
#       Genus = rect["genus"])
#   }
#   else{e_taxon <- e_taxon %>%  add_row(Phylum = NA)}
# }
# higher_tax <- e_taxon %>% 
#   mutate(`Sample ID` = submit$`PTM#`)
# 
# e_taxon <- tax[FALSE,]
```

## Voucher Sheet
** CAUTION CHECK THROUGH THIS PART**
Edit the data to match the formatting of BOLD data
```{r}
c_info <- submit %>% 
  mutate(# Voucher
    Sample_ID = paste0("PTM",`PTM#`),
    Field_ID = Sample_ID,
    Museum_ID = `Accession #`,
    Collection_Code = "",
    Insitution_Storing = "University of British Columbia, Herbarium",
    #Taxon
    Species = `Determination in the field`,
    Identifier = `Determined by`,
    Identifier_Email = "",
    Identifier_Insitution  = "University of British Columbia",
    `Identification_Method` = "",
    `Taxonomy Notes` = "",
    #collect
    Collectors = "",
    Collection_Date = `Date Collected`,
    Country_Ocean = "Canada", # depends on data set
    State_Province = "British Columbia", #depends on dataset
    Sector = "",
    Region = "",
    #Specimen Details
    Sex = "",
    Reproduction = `Reproductive Status`,
    Life_Stage = "",
    Extra_info = "",
    Note = "",
    Voucher_stat = "",
    Tissue_des = "",
    Assoc_tax = "",
    Assoc_spe = "",
    External = "")%>%
  separate(Locality, c("Sector","Exact_site"),",")

c_info <- c_info %>% 
  mutate(Elevation = "",
         Depth = "",
         Elevation_Precision = "",
         Depth_Precision = "",
         GPS_Source	= "",
         Coordinate_Accuracy= "",
         Event_Time = "",
         Collection_Date_Accuracy	= "",
         Habitat = c_info$Habitat,
         Sampling_Protocol = "",
         Collection_Notes = "",
         Site_Code = "",
         Collection_Event_ID = "")
```

```{r}
#selecting the relevant columns
voucher <- c_info %>%
  select(`Sample_ID`,`Field_ID`, `Museum_ID`, `Collection_Code`, `Insitution_Storing`)

names(voucher) <- names(e_info)
```

```{r}
s <- c_info %>% 
  select(`Sample_ID`)

t_high <-  higher_tax %>%
  select(`Phylum`, `Class`, `Order`, `Family`, `Subfamily`, `Genus`)
t_sub <- c_info %>% 
  select(`Species`,`Identifier`, `Identifier_Email`, `Identifier_Insitution`, `Identification_Method`, `Taxonomy Notes`)

taxon <- cbind(s,t_high,t_sub)
names(taxon) <- names(e_taxon)

collect <-  c_info %>%
  select(`Sample_ID`,`Collectors`, `Collection_Date`, `Country_Ocean`,`State_Province`, `Region`, `Sector`, `Exact_site`, `Latitude`, `Longitude`,`Elevation`, Depth, Elevation_Precision, Depth_Precision, Depth_Precision, GPS_Source, Coordinate_Accuracy, Event_Time, Collection_Date_Accuracy, Habitat, Sampling_Protocol, Collection_Notes, Site_Code, Collection_Event_ID)

names(collect) <- names(e_collect)
```

```{r}
sd <- c_info %>% 
  select(Sample_ID,Sex, Reproduction, Life_Stage, Extra_info, Note, Voucher_stat, Tissue_des, Assoc_tax, Assoc_spe, External)

names(sd) <- names(e_specimen)
```

# Finally saving your data
- always look through the final records to see if the information has been placed correctly

!!! Look into writexl
https://github.com/ropensci/writexl

```{r}
lod <- list("Voucher Info" = voucher, "Taxonomy" = taxon, "Specimen Details" =sd, "Collection Data" = collect )
write.xlsx(lod, file = paste0("results/",filename,".xlsx"), append = TRUE)
```

## beeps when done
```{r}
beep(sound = 5)
```

