---
title: "merge_bold_pub"
author: "Jasmine"
date: '2019-02-26'
output: html_document
---

Make sure you have all the libraries installed
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
library(here)
```

## Read in your files
```{r}
bold <- read_xlsx(here("03_publication/examples/bold.xlsx"),skip = 2)
s1 <- read_csv(here("03_publication/examples/TableS1_V2.csv"))

s1 <- filter(s1,!is.na(s1$`Collection Number (PTM)`))
```

## Fixing up the fields before joining
note this is done specifically with the TableS1_V2.csv. Some modification to the columns used may vary
```{r}
acc <- bold %>% mutate(PTM = stringr::str_replace(bold$`Field ID`,"PTM", "")) %>% select(PTM,`PSBA Accession`,`COI-5P Accession`,`rbcL Accession`) %>% 
  transform(PTM = as.numeric(PTM))
```


## Populating the Genbank number field
Note: this keeps the individual columns for each gene as well --> do a selection to remove them
```{r}
s1_acc <- left_join(s1, acc, by = c("Collection Number (PTM)" = "PTM"))

s1_na <- replace(s1_acc, is.na(s1_acc), "")

pub <- s1_na %>% 
  mutate(`Genbank Number` = paste(s1_na$PSBA.Accession,s1_na$COI.5P.Accession,s1_na$rbcL.Accession))
```

## Saving

The final file will be in the results folder
```{r}
write.csv(pub, file ="results/TableS1_V2_genbank.csv")
```